
\setcounter{chapter}{-1}
\chapter{Introduction to the Course}

\begin{itemize}
    \item Textbook: \emph{Computer Organization and Architecture}, 10th ed., William Stallings
    \item Grading Policy
    \begin{itemize}
        \item Assignment: 4 20\%
        \item Term Project: Report 15\% \\
            might not happen and be changed to attendance
        \item Midterm and Final Exam 30\% or 35\%
    \end{itemize}
    \item Covered Content
    \begin{enumerate}
        \item \textbf{Overview}
        \begin{enumerate}[label = Ch. \arabic*:, labelindent=0pt, itemindent=1em, leftmargin=2em]
            \setcounter{enumii}{0}
            \item Introduction
            \item Computer Evolution and Performance
        \end{enumerate}
        \item \textbf{The Computer System}
        \begin{enumerate}[label = Ch. \arabic*:, labelindent=0pt, itemindent=1em, leftmargin=2em]
            \setcounter{enumii}{2}
            \item Top-Level View of Computer Functions
            \item Cache Memory
            \item Internal Memory
            \item External Memory
            \item Input / Output
        \end{enumerate}
        \item Arithmetic and Logic \\
            Skipped
        \item \textbf{The Central Processing Unit}
        \begin{enumerate}[label = Ch. \arabic*:, labelindent=0pt, itemindent=1em, leftmargin=2em]
            \setcounter{enumii}{11}
            \item Instruction Sets: Characteristics and Functions
            \item Instruction Sets: Addressing Modes and Formats
            \item Processor Structure and Function
            \item Reduced Instruction Set Computer
        \end{enumerate}
        \item The Parrallel Organization
        \begin{enumerate}[label = Ch. \arabic*:, labelindent=0pt, itemindent=1em, leftmargin=2em]
            \setcounter{enumii}{16}
            \item Parrallel Processing
            \item Multicore Computers
        \end{enumerate}
        \item The Control Unit 19--1
        \begin{enumerate}[label = Ch. \arabic*:, labelindent=0pt, itemindent=1em, leftmargin=2em]
            \setcounter{enumii}{18}
            \item Control Unit Operation 19--1
            \item Microprogrammed Control 20--1
        \end{enumerate}
    \end{enumerate}
\end{itemize}


\chapter{Introduction}

\section{Computer Architecture vs. Organization}

\subsection*{Computer Architecture}

\begin{itemize}
    \item Attributes of a system visible to the programmer
    \begin{itemize}
        \item Instruction set
        \item Number of bits used to represent various datatypes
        \item I/O mechanims
        \item Techniques for addressing memory
        \item etc.
    \end{itemize}
    \item Have a direct impact on the logical execution of a program
\end{itemize}

\subsection*{Computer Organization}

\begin{itemize}
    \item The operation units and their interconnections that realize the architectural specifications
    \begin{itemize}
        \item Hardware details transparent to the programmer
        \item Control signals
        \item Interfaces between the computer and the peripherals
        \item Memory technology used
        \item etc.
    \end{itemize}
\end{itemize}




